### 2020-06-05 chapters/background.tex
Test

Batch processing is the processing of data grouped by batches. A batch is a collection of data points that have been grouped by certain criteria. Examples of such criteria could be \textit{"all the transactions from the past two months"}, \textit{"the last million transactions"} or even \textit{"all the transactions made by an user"}. Batch data sets are known as static data \cite{Martin-Batch-Defin}. An example application of such a technique would be processing all the keywords searched for in the google search engine. For instance, grouping the results by keyword and counting how many times a keyword was searched for trend analysis purposes. Batch processing is the oldest paradigm between batch and stream processing. Historically speaking, Apache Hadoop \cite{hadoop} was one of the first widely known technologies for large-scale batch processing, created in 2005. Nearly a decade later, Apache Spark \cite{spark} was released. Spark is capable of doing not only on-disk processing like Hadoop but also in-memory processing, achieving lower processing times. As more and more data became available, the need for stream processing increased. Stream processing refers to the processing of data streams. A data stream is essentially an unbounded data set. This is in contrast to batch processing which works on a finite data set. This distinction is important since some computations are impossible to compute accurately on unbounded data sets. Such an example of an impossible computation on an infinite data set is grouping incoming data by an ID and counting distinct elements.

@@

% streaming as a proper superset of batch
Distinguishing between batch and stream processing is important to understand how a monitoring system on top of the latter would work. The Venn diagram \ref{fig:stream-superset} was taken from \cite{batch-is-a-special-case-of-streaming} and illustrates the relation of both processing techniques. A data stream is unbounded or infinite while a bounded data stream has a beginning and an end. Using these definitions, the content of an unbounded data stream contains that of a bounded data set. That is, from an infinite stream, it is possible to fix a left and right boundary in order to obtain a bounded and finite data set. This relation is visible in the \ref{fig:unbounded-bounded-superset} diagram from \cite{batch-is-a-special-case-of-streaming}.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.7]{figures/streaming-subset-batch.png}
      \caption{Stream processing as a superset of batch processing}
      \label{fig:stream-superset}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.7]{figures/unbounded-superset-bounded.png}
      \caption{Unbounded data streams as a superset of bounded data sets}
      \label{fig:unbounded-bounded-superset}
    \end{center}
\end{figure}

@@

Recalculate-From-Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each window update. For instance, the computation of a sum over a sliding window using RFS means that whenever the window changes all elements are summed. This approach works for every aggregation. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing.

Subtract-On-Evict (SOE) on the other hand, is a SWAG algorithm that incrementally computes the aggregation value. For each element insertion or eviction from the window, the aggregation value is updated, in contrast to RFS which would recompute the entire aggregation. SOE relies on expiring the evicted elements from the aggregation state. Hence, SOE only works for \textit{invertible} aggregations. To illustrate, consider the computation of a sum aggregation over a sliding window. The total sum of the window will be the aggregation state or value. When inserting an element, the updated total window sum equals the total sum plus the new element. The sum aggregation is considered \textit{invertible} because, when evicting an element, the new aggregation state is the total sum computed this far minus the evicted element. With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable.

@@

>>>>

Batch processing is the processing of data grouped by batches. A batch is a collection of data points that have been grouped by certain criteria. Examples of such criteria could be \textit{"all the transactions from the past two months"}, \textit{"the last million transactions"} or even \textit{"all the transactions made by an user"}. Batch data sets are known as static data \cite{Martin-Batch-Defin}. An example application of such a technique would be processing all the keywords searched for in the google search engine. For instance, grouping the results by keyword and counting how many times a keyword was searched for trend analysis purposes. Batch processing is the oldest paradigm between batch and stream processing. Historically speaking, Apache Hadoop \cite{hadoop} was one of the first widely known technologies for large-scale batch processing, created in 2005. Nearly a decade later, Apache Spark \cite{spark} was released. Spark is capable of doing not only on-disk processing like Hadoop but also in-memory processing, achieving lower processing times. As more and more data became available in the data stream format, the need for stream processing increased. Stream processing refers to the processing of data streams. A data stream is essentially an unbounded data set. This is in contrast to batch processing which works on a finite data set. This distinction is important since some computations are impossible to compute accurately on unbounded data sets. Such an example of an impossible computation on an infinite data set is grouping incoming data by an ID and counting distinct elements.

@@

% streaming as a proper superset of batch
Distinguishing between batch and stream processing is important to understand how a monitoring system on top of the latter would work. The Venn diagram \ref{fig:stream-superset} was taken from \cite{batch-is-a-special-case-of-streaming} and illustrates the relation of both processing techniques. A data stream is unbounded or infinite while a bounded data stream has a beginning and an end. Using these definitions, the content of an unbounded data stream contains that of a bounded data set. That is, from an infinite stream, it is possible to fix a left and right boundary in order to obtain a bounded and finite data set. This relation is visible in Figure \ref{fig:unbounded-bounded-superset} from \cite{batch-is-a-special-case-of-streaming}.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/streaming-subset-batch.png}
      \caption{Stream processing as a superset of batch processing}
      \label{fig:stream-superset}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/unbounded-superset-bounded.png}
      \caption{Unbounded data streams as a superset of bounded data sets}
      \label{fig:unbounded-bounded-superset}
    \end{center}
\end{figure}

@@

Recalculate-From-Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each update. For instance, the computation of a sum over a sliding window using RFS means that whenever the window changes all elements are summed. This approach works for every aggregation. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing. Figure \ref{fig:pseudo-rfs} shows pseudocode for RFS \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms}.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.45]{figures/rfs.png}
      \caption{Recalculate-From-Scratch algorithm}
      \label{fig:pseudo-rfs}
    \end{center}
\end{figure}

Subtract-On-Evict (SOE) on the other hand, is a SWAG algorithm that incrementally computes the aggregation value. For each element insertion or eviction from the window, the aggregation value is updated, in contrast to RFS which would recompute the entire aggregation. SOE relies on expiring the evicted elements from the aggregation state. Hence, SOE only works for \textit{invertible} aggregations. To illustrate, consider the computation of a sum aggregation over a sliding window. The total sum of the window will be the aggregation state or value. When inserting an element, the updated total window sum equals the total sum plus the new element. The sum aggregation is considered \textit{invertible} because, when evicting an element, the new aggregation state is the total sum computed this far minus the evicted element. With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable. Figure \ref{fig:pseudo-soe} shows pseudocode for SOE \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms}.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.45]{figures/soe.png}
      \caption{Subtract-On-Evict algorithm}
      \label{fig:pseudo-soe}
    \end{center}
\end{figure}

@@

%\section{Aggregations on data types}

%Some aggregations are only possible on certain data types. For instance, computing an average on a set of strings is not possible. In this Section we present the data types we will work with and the set of possible aggregations to compute on them.

%We consider four data types: \textit{numericals}, \textit{strings}, \textit{dates} and \textit{categoricals}. Numerical variables can be measured. When compared to classical programming data types, for example those of the C programming language, numericals comprise variable types \textit{short}, \textit{long}, \textit{int}, \textit{float}, \textit{double} and valid combinations of the previous --- \textit{e.g.} \textit{long int}. On the other hand, Strings are sequences of characters. The \textit{string} type is the most generic data type in that any other type can be converted to a string. Every variable, for example \textit{x}=3, can be turned into a string \textit{s}, for example \textit{s}="3". Similarly, a date is a sequence of characters and thus a string. However, despite being a string, it has semantics unlike a regular one. Knowing the date format, one can extract the day, month and year. For that reason we make the distinction between the \textit{string} and \textit{date} data types. Lastly, a \textit{categorical} is a variable that takes one of a relatively small and known set of possible values. A variable is assigned to a particular category based on its properties. The categorical variable is also a sequence of characters and therefore a string variable. The main difference between the categorical data type and the string is that we know the set of possible labels the categorical variable might take and this affects the type of computable aggregations.

%On a set of numerical variables, we can compute sums, minima, maxima, averages, standard deviations and counts (of zeros, infinities and others). When operating on a set of strings, we can aggregate the collection into a concatenation of strings, a count, a membership query, cardinality.

\section{Exponential Moving Averages}
Both SOE and RFS require saving the entire window contents. IS THERE A WAY WE CAN EMULATE AN WINDOW....

### 2020-06-05 chapters/introduction.tex
Aggregation on data types
Test

\section{Context} \label{sec:context}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. Different use cases process data in different ways for different purposes. For instance, e-commerce platforms need to process a large number of transactions, for example, those made daily by all Alibaba's clients while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube and Netflix, try to guarantee that media content reaches up to millions of users simultaneously. Furthermore, this media content may have been previously recorded and stored, or may be generated on the fly. Social networks are also responsible for generating large volumes of data. A social network like Twitter, for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day \cite{twitter-throughput}. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. These need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the protected system.

@@

Both in batch and stream processing, we usually want to apply operations on data, for example, to compute the maximum value, an average or a count of distinct elements. Operations applied to data sets that produce a single result are commonly known as aggregate operations. In stream processing, we apply these aggregations to data streams to obtain valuable information. However, since a data stream is, effectively speaking, an unbounded data set, if our aggregation requires a finite domain, there is the need to apply windowing techniques. A thorough analysis to stream windowing techniques is done in the Background Section of this document but for now it suffices to say that windowing transforms an unbounded data stream into a finite data set by defining a time or tuple-based window size --- \textit{i.e.} events from the last three days or the last $10^6$ events, respectively.

@@

Many market solutions monitor streams of credit card transactions with the intent of fraud detection. In the context of this Thesis, we will work side by side with Feedzai. Feedzai offers fraud prevention solutions for the previously presented use case of monitoring an infinite stream of credit card transactions and detecting fraud. We will work side by side with them, taking advantage of the multiple real-life data sets from the financial fraud space they can provide.

@@

In Chapter \ref{chap:approach} we define the scope of this project, the research questions we intend to answer, our experimental methodology and our future planning.

>>>>

\section{Context} \label{sec:context}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. Different use cases process data in different ways for different purposes. For instance, e-commerce platforms need to process a large number of transactions, for example, those made daily by all Alibaba's clients, while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube and Netflix, try to guarantee that media content reaches up to millions of users simultaneously. Furthermore, this media content may have been previously recorded and stored, or may be generated on the fly. Social networks are also responsible for generating large volumes of data. A social network like Twitter, for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day \cite{twitter-throughput}. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. These need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the protected system.

@@

Both in batch and stream processing, we usually want to apply operations on data, for example, to compute the maximum value, an average or a count of distinct elements. Operations applied to data sets that produce a single result are commonly known as aggregate operations. In stream processing, we apply these aggregations to data streams to obtain valuable information. However, since a data stream is, effectively speaking, an unbounded data set, if our aggregation requires a finite domain, there is the need to apply windowing techniques. A thorough analysis to stream windowing techniques is done in the Background Section of this document but for now it suffices to say that windowing transforms an unbounded data stream into a finite data set by defining a time or tuple-based window size --- \textit{i.e.} events from the last three days or the last million events, respectively.

@@

Many market solutions monitor streams of credit card transactions with the intent of fraud detection. In the context of this Thesis, we will work side by side with Feedzai. Feedzai offers fraud prevention solutions for the previously presented use case of monitoring an infinite stream of credit card transactions and detecting fraud. We will work with them, taking advantage of the multiple real-life data sets they can provide from the financial fraud space.

@@

%In Chapter \ref{chap:approach} we define the scope of this project, the research questions we intend to answer, our experimental methodology and our future planning.
